---
title: "PHP2560 SEIR Project"
author: "Breanna Richards, Tim Hedspeth, Nancy Liu"
date: "11/26/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
library(latex2exp)
```

Our Model:

- Some exposed individuals will enter the quarantined class and others enter the infectives.

- Individuals in quarantined and infected classes will enter diagnosed class

- Note that not all infected people will be diagnosed

- They will either enter the recovered class or die

- $\epsilon$ rate of exposed people moving into infective class

- $\lambda$ rate of exposed people moving into quarantined class

- $\sigma$ rate of quarantined people moving into diagnosed class

- $\theta$ rate of infectives moving into diagnosed class

- $\gamma$ rate of diagnosed moving into recovered class

- $\delta$ rate of infected and diagnosed people dying

- $\beta(t)$ transmission rate



Default Rate Values:

- $\epsilon = (1/3)(2/5)$

- $\lambda = (1/3)(3/5)$

- $\delta = (15/100)(1/21)$

- $\theta = 1/3$

- $\sigma = 1/3$

- $\gamma = 1/21$

- $k = 0.1$

Infection rate $\beta(t)$ is most important parameter because the model is most sensitive to it. 

Default Initial Number of People: 
- Exposed: E(0) = 477
- Infectives: I(0) = 286
- Quarantined: Q(0) = 191
- Diagnosed: J(0) = 848
- Recovered: R(0) = 1213


This is an example of our simulations and the plots!  

## Starting the Simulation


```{r}



get_diagnosed <- function( e = 477, i = 286, q = 191, 
                j = 848, r = 1213, epsilon = (1/3)*(2/5),
lambda = (1/3)*(3/5), delta = (15/100)*(1/21), 
theta = 1/3, sigma = 1/3, gamma = 1/21, k = 0.1, b = (31 + t)/(22 + 5*t), max_e = NULL,
max_i = NULL, max_q = NULL, max_j = NULL) {

days <- 400

if (!is.null(max_e)) {
e <- runif(1, 1, max_e)
}

if (!is.null(max_i)) {
i <- runif(1, 1, max_i)
}

if (!is.null(max_q)) {
q <- runif(1, 1, max_q)
}

if (!is.null(max_j)) {
j <- runif(1, 1, max_j)
}

if (!is.null(max_r)){
r <- runif(1, 1, max_r)
}


for (t in 1:(days - 1)) {


e[t+1] <- e[t] + b*((k*e[t])+i[t]) - (epsilon+lambda)*e[t]
i[t+1] <- i[t] + epsilon*e[t] - (delta + theta)*i[t]
q[t+1] <- q[t] + lambda*e[t] - sigma*q[t]
j[t+1] <- j[t] + theta*i[t] + sigma*q[t] - (delta + gamma)*j[t]
r[t+1] <- r[t] + gamma*j[t]


if (b == (31 + t)/(22 + 5*t)) {
  
  b <- (31 + (t+1))/(22 + 5*(t+1))
  
}



}

all_data <- as.data.frame(cbind(Time = 1:days, Exposed = e, Infectives = i, 
                                Quarantined = q, Diagnosed = j, Recovered = r))

return(all_data)
}


get_diagnosed()
example_data <- get_diagnosed()

```

This is a preview of what our data looks like! Users will be able to pick what columns they want to look at. 

```{r}

head(example_data)

```


## Visualization Examples (RESULTS)

```{r}

df_full <- example_data %>%
pivot_longer(!Time, names_to = "people_type", values_to = "num_people")

```


A checkbox will indicate which lines we want to see!

```{r}

# plot separately
ggplot(df_full[df_full$people_type == "Diagnosed", ], aes(x = Time,
                              y = num_people)) + geom_line() 


ggplot(df[df$people_type == "Exposed", ], aes(x = Time, 
                            y = num_people)) + geom_line() +
  labs(title = TeX("Number of Exposed Cases with $\\beta = 0.6781$"))

ggplot(df[df$people_type == "Infectives", ], aes(x = Time, 
                          y = num_people)) + geom_line() +
  labs(title = TeX("Number of Infectives with $\\beta = 0.6781$"))

ggplot(df[df$people_type == "Quarantined", ], aes(x = Time, 
                            y = num_people)) + geom_line() +
  labs(title = TeX("Number of Quarantined Cases with $\\beta = 0.6781$"))

ggplot(df[df$people_type == "Recovered", ], aes(x = Time, 
                            y = num_people)) + geom_line() +
  labs(title = TeX("Number of Recovered Cases with $\\beta = 0.6781$"))


ggplot(df[df$people_type ==  "Diagnosed" | df$people_type == "Recovered", ],
       aes(x = Time, y = num_people)) + geom_line(aes(col = people_type)) +
  labs(title = TeX("Number of Diagnosed and Recovered Cases with $\\beta(t)$"))

ggplot(df, aes(x = Time, y = num_people)) + geom_line(aes(col = people_type)) +
  labs(title = TeX("Number of Cases by People Type with $\\beta(t)$"))


```


# Animate

```{r}

library(plotly)
library(gganimate)



# button that says run the simulation

# function to run visual 
run_animation <- function(df_use) {

df_use <- accumulate_by(df_use, ~ Time)

df_use %>%
  plot_ly(
    x = ~ Time,
    y = ~ num_people,
    split = ~ people_type,
    frame = ~ frame,
    type = 'scatter',
    text = ~ num_people,
    mode = 'lines',
    hoverinfo = "text",
    line = list(simplyfy = F)
  ) %>%
  add_markers(ids = ~ num_people) %>%
  layout (
    xaxis = list(
      title = "Day",
     # range = c()
      zeroline = F
    ),
    yaxis = list(
      title = "Number of People",
      zeroline = F
    )
  ) %>%
  animation_opts(
    frame = 200,
    transition = 0,
    redraw = T
  ) %>%
  animation_slider(
    hide = F
  ) %>%
  animation_button(
    x = 1, xanchor = "right", y = 0, yanchor = "bottom"
  )
}


run_animation(df_full)

```


 
 
 
 
```{r} 

 # unused code
# p +
#   
#   gganimate::transition_reveal(Time) +
#   
#   labs(title = "{frame_along}: Number of Diagnosed Cases with beta(t)",
#        caption = "diagnosed over time")


# p <- ggplot(df[df$people_type == "Diagnosed", ], aes(x = Time,
#                               y = num_people)) + geom_line() +
#   labs(title = TeX("Number of Diagnosed Cases with beta(t)")) +
#   theme_bw()
# 
# q <- ggplot(df, aes(x = Time, y = num_people)) + geom_line(aes(col = people_type)) +
#   labs(title = "Number of Cases by People Type with beta(t)")
# 
# 
# fig <- ggplotly(p, tooltip = "text")
# fig
# 
# fig <- fig %>%
#   animation_slider(
#     currentvalue = list(prefix = "Day ", font = list(color="red")))

```

https://www.datanovia.com/en/blog/gganimate-how-to-create-plots-with-beautiful-animation-in-r/


